Deciding Bitvector Termination with Program Synthesis

### 2014-07-02 papers/termination/paper.tex
revising sec 6
Nudge introduction, introduce the halting problem

Proving program termination is typically done by finding a \emph{ranking function}
for the program states, i.e. a map from the program's state space to a well-ordered set.
%% As this definition of a ranking function is very general, research is often limited to some
%% convenient and tractable form of ranking functions, most frequently \emph{linear ranking functions} (see Section~\ref{sec:ranking.functions}). 
%% In doing so, an analysis ensures that such a ranking function will be found if one  exists, 
%% while failing to prove termination for terminating programs whose ranking functions fall short of the considered restriction.

@@

If the initial states $I$ are ignored, this loop cannot be shown to terminate (e.g. any state with $y = 0$ would lead to a
non-terminating execution).

However, formula {\bf [CT]} is satisfiable, as witnessed by the ranking function
$R(\langle x,y\rangle ) = x$ and supporting invariant $W(\langle x, y \rangle ) \trangleq y = 1$.\\

>>>>

The halting problem has been of central interest to computer scientists since it was first
considered by Turing in 1936~\cite{turing}.  Informally, we can say that the halting problem
is concerned with answering the question ``does this
program run forever, or will it eventually terminate?''

Proving program termination is typically done by finding a \emph{ranking function}
for the program states, i.e. a monotone map from the program's state space to a well-ordered set.
Historically, the search for ranking functions has been constrained in various syntactic ways, leading to
incompleteness, and
is performed over abstractions that do not soundly capture the behaviour of physical computers.
In this paper, we present a sound and complete method for deciding whether a program terminates.

%% As this definition of a ranking function is very general, research is often limited to some
%% convenient and tractable form of ranking functions, most frequently \emph{linear ranking functions} (see Section~\ref{sec:ranking.functions}). 
%% In doing so, an analysis ensures that such a ranking function will be found if one  exists, 
%% while failing to prove termination for terminating programs whose ranking functions fall short of the considered restriction.

@@

If the initial states $I$ are ignored this loop cannot be shown to terminate, since any state with $y = 0$ and $x > 0$
would lead to a non-terminating execution.

However, formula {\bf [CT]} is satisfiable, as witnessed by the ranking function
$R(\langle x,y\rangle ) = x$ and supporting invariant $W(\langle x, y \rangle ) \triangleq y = 1$.
This constitutes a proof that the program as whole terminates, since the loop always begins
executing in a state that guarantees its termination.
